---
title: "Cyclystic Data analysis capstone proyect"
author: "Christian Calvanese"
date: "22, July 2022"
output: 
  html_document:default
  pdf_document:default
  word_document:default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Background of the dataset company

In 2016, Cyclistic launched a successful bike-share offering. Since
then, the program has grown to a fleet of 5,824 bicycles that are
geotracked and locked into a network of 692 stations across Chicago. The
bikes can be unlocked from one station and returned to any other station
in the system anytime.

There are two types of cyclist, those who purchase casual tickets and
those who purchase annual memberships. The marketing team believes that
maximizing the number of annual members will be key to future growth.
Rather than creating a marketing campaign that targets all-new
customers, there is a very good chance to convert casual riders into
members.

# Installing requiered packages

```{r}
install.packages("tidyverse", "janitor", "lubridate")
library(tidyverse) 
library(janitor) 
library(lubridate)
install.packages("janitor")
```

After the package installation we proceed to merge the data of a full year so we can clean it later. 
This is the data source:
##Data Source: \## <http://divvy-tripdata.s3.amazonaws.com/index.html>


```{r}
m1 <-read.csv("./cyclistic_data/202101-divvy-tripdata.csv") 
m2 <-read.csv("./cyclistic_data/202102-divvy-tripdata.csv") 
m3 <-read.csv("./cyclistic_data/202103-divvy-tripdata.csv") 
m4 <-read.csv("./cyclistic_data/202104-divvy-tripdata.csv") 
m5 <-read.csv("./cyclistic_data/202105-divvy-tripdata.csv")
m6 <-read.csv("./cyclistic_data/202106-divvy-tripdata.csv") 
m7 <-read.csv("./cyclistic_data/202107-divvy-tripdata.csv") 
m8 <-read.csv("./cyclistic_data/202108-divvy-tripdata.csv") 
m9 <-read.csv("./cyclistic_data/202109-divvy-tripdata.csv") 
m10 <-read.csv("./cyclistic_data/202110-divvy-tripdata.csv") 
m11 <-read.csv("./cyclistic_data/202111-divvy-tripdata.csv") 
m12 <-read.csv("./cyclistic_data/202112-divvy-tripdata.csv")
```

After we assign each month to a variable we combine and check for empty rows or columns,
we remove them and check if they are significative to the dataset total datapoints.


```{r}
bike_rides <- rbind(m1,m2,m3,m4,m5,m6,m7,m8,m9,m10,m11,m12)
dim(bike_rides) 
bike_rides <- janitor::remove_empty(bike_rides, which = c("cols")) 
bike_rides <- janitor::remove_empty(bike_rides, which =c("rows")) 
dim(bike_rides)
```

Now that we have our merged data we check if the column datapoints are in line with
the datatype, if they are not we parse them.

```{r}
##parsing Strings to POSIXcl
bike_rides$started_at <- ymd_hms(bike_rides$started_at)
bike_rides$ended_at <- ymd_hms(bike_rides$ended_at)
```

POSIXcl is a date format in R.

```{r}
##calculating difference of time in minutes
bike_rides$travel_time <- difftime(bike_rides$ended_at,
bike_rides$started_at, units = "mins") %>%  round 

##getting weekdays 
bike_rides$day_of_ride <- weekdays(bike_rides$started_at)
```

Last but not least we format the station IDs and we save it for Tableau viz.

```{r}
##formating the stations IDs from STRING to Double
bike_rides$start_station_id <- as.double(bike_rides$start_station_id)
bike_rides$end_station_id <- as.double(bike_rides$end_station_id)

##saving the data frame to work with tableau 
save(bike_rides,file="bike_rides.Rda")
```



